Statistical methods for relative quantification of post-translational modifications in global proteomics experiments

### 2021-09-29 supplementary/ptm_sm.tex
Made updates to both main and supp

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Simulation 1}

@@

\subsubsection{Simulation 2}

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Benchmark experiment}
\label{sec:benchmark}

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Datasets : Biological investigation}

\subsection{IpaH7.8}
\label{sec:ipah}

@@

The results of the proposed method to this experiment can be seen in \sfigref{fig:ipah_figures}.

@@

\clearpage
\subsection{Shigella flexneri}
\label{sec:shigella}

@@

The results of the proposed method to this experiment can be seen in \sfigref{fig:No_Diff_Shigella_PTM} and \sfigref{fig:Diff_Shigella_PTM}.

@@

\subsection{USP30}
\label{sec:usp30}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{sim_new/USP30_profile_plot}
\caption{Comparing the global profiling of protein $P52209$ with the modification of the protein at site $K059$. The modification appears generally unchanged between all conditions, whereas the global profiling run shows the CCCP and Combo conditions have a higher relative abundance compared to the Control and USP30\_OE. This indicates that the modification actually had a major effect when comparing CCCP and Combo to Control and USP30\_OE. This would have been missed without adjusting for global protein changes.}
\label{fig:USP30_profile_plot}
\end{figure}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample size calculation and power analysis}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Computer simulation 1 - label free}

@@

\subsubsection{Computer simulation 2 - missing values and low replicates}

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SpikeIn benchmark - KGG enriched}
\label{sec:benchmark}

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Datasets : Biological investigation}
\label{sec:bio_investigations}

\subsection{Human-1mix-TMT - ubiquitination}
\label{sec:ipah}

@@

The results of the proposed method to this experiment can be seen in \sfigref{fig:ipha_venn_d} and \sfigref{fig:ipah_figures}.

\begin{figure}[h!]
\centering
\includegraphics[height=.7\textwidth]{sim_new/ipah_venn_diagramm.png}
\caption{The overlap of differencial modified peptides between the PTM model with and without global protein level adjustment. More PTMs became insignificant after adjustment then became significant. This indicates that for the peptides that became insignificant in the adjusted model, their change in abundance was mainly due to changes in global protein abundance. An additional question that must be addressed is if the decrease in significant peptides is due to the increased variance that comes from adjustment. This was tested by looking for modified peptides who's adjusted log fold change was within 10\% of the unadjusted log fold change but became insignificant after adjustment. When this test was applied on this experiment, only one peptide became insignificant due to an increase in variance. Thus we can conclude that the drop off in significant peptides was truly due to changes in global protein abundance.}
\label{fig:ipha_venn_d}
\end{figure}

@@

\clearpage
\subsection{Mouse-2mix-TMT - phosphorylation}
\label{sec:shigella}

@@

The results of the proposed method to this experiment can be seen in \sfigref{fig:shig_venn_d},  \sfigref{fig:No_Diff_Shigella_PTM}, and \sfigref{fig:Diff_Shigella_PTM}.

\begin{figure}[h!]
\centering
\includegraphics[height=.7\textwidth]{sim_new/shig_venn_diagramm.png}
\caption{The overlap of differencial modified peptides between the PTM model with and without global protein level adjustment. Much like in \sfigref{fig:ipha_venn_d}, more PTMs became insignificant after adjustment then became significant. This indicates that for the peptides that became insignificant in the adjusted model, their change in abundance was mainly due to changes in global protein abundance. Again we tested if the decrease in significant peptides was due to the increased variance that comes from adjustment. This was tested by looking for modified peptides who's adjusted log fold change was within 10\% of the unadjusted log fold change but became insignificant after adjustment. When this test was applied on this experiment, 776 peptides became insignificant due to an increase in variance. This is higher than in \sfigref{fig:ipha_venn_d}, however it is still a very small portion (4\%) of the peptides that became insignificant after adjustment. Thus we can conclude that the drop off in significant peptides was mainly due to changes in global protein abundance.}
\label{fig:shig_venn_d}
\end{figure}

@@

\subsection{Human label free quantification - no global profiling run}
\label{sec:usp30}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{sim_new/USP30_profile_plot}
\caption{Comparing the global profiling of protein $P52209$ with the modification of the protein at site $K059$. The modification appears generally unchanged between all conditions, whereas the global profiling run shows the CCCP and Combo conditions have a higher relative abundance compared to the Control and USP30\_OE. This indicates that the modification actually had a major effect when comparing CCCP and Combo to Control and USP30\_OE. This would have been missed without adjusting for global protein changes.}
\label{fig:USP30_profile_plot}
\end{figure}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample size calculation and power analysis}
\todo{Devon: Update this section per Olga's commentsv}
